Evolution of Processes in Delivering Reliable Software in the European Scientific Arena

### 2018-06-26 paper-ieee-trans.tex
paragraph break, check reviewer 1 comments
fix two issues sent by Eli

\IEEEPARstart{D}{uring} the last 15 years the provision of an European e--Infrastructure,
supporting unified access to large--scale computing and intense data analysis
applications, has been driven by the needs of scientific communities as part of their
pan--European research collaborations. One such example is the Worldwide
Large Hadron Collider (LHC) Computing Grid infrastructure, driven
by the particle physics scientific community, built to satisfy the computing needs
of processing, storing and analysis of the very large amount of data produced by the
LHC particle accelerator. A significant number of innovation projects and technology transfer
activities -- later discussed in Section~\ref{sec:ev} -- from the European Commission
(EC)-funded 5th, 6th and 7th Framework Programmes, and recently from the Horizon 2020 programme
\cite{h2020} had as their main goal to develop software solutions that provide a seamless and
transparent access by researchers to distributed computing and data resources.

@@

The EC projects considered in this paper are listed chronologically in Table~\ref{tab:eup}.
Each project paved the way for the following ones, according to a logic evolution in which
the role of the quality and reliability of the software has
consistently increased. As shown in Table~\ref{tab:feat}, the most recent projects are more
committed to and aware of software reliability practices.
Table~\ref{tab:feat} also shows the topics addressed in the following sections, highlighting
the commonalities and specific features on software reliability.

@@

The CI pipeline is composed from a set of required checks, their exit status
is reported back to GitHub. The changes can be prevented if those tests are
not successfully executed.
As each change is validated, the chances
of early detection of defects increases. Within this scenario, the cost of defect solving
is dramatically reduced and the reliability of the software solutions improved, as any
bug or design issue is likely to be detected and subsequently corrected in this phase.

@@

\subsubsection{Continuous Delivery}
As DevOps suggests, frequent releases positively affect the reliability of the software as
they reduce the time in which a given defect is exposed or has an impact, allowing the 
development teams to act promptly based on the regular feedback \cite{chen2015}. The software updates of
{\sl INDIGO--DataCloud} products, that have been taking place since the second major release, are passing through
a Continuous Delivery (CD) pipeline that adds the packaging of the software right after the
successful execution of the CI pipeline previously described. Consequently, CI--validated
software components are automatically delivered in online repositories. Nonetheless,
expert supervision to validate the results, is required.

>>>>

\IEEEPARstart{D}{uring} the last 15 years the provision of an European e--Infrastructure,
supporting unified access to large--scale computing and intense data analysis
applications, has been driven by the needs of scientific communities as part of their
pan--European research collaborations. One such example is the Worldwide
Large Hadron Collider (LHC) Computing Grid infrastructure, driven
by the particle physics scientific community, built to satisfy the computing needs
of processing, storing and analysis of the very large amount of data produced by the
LHC particle accelerator.

A significant number of innovation projects and technology transfer
activities -- later discussed in Section~\ref{sec:ev} -- from the European Commission
(EC)-funded 5th, 6th and 7th Framework Programmes, and recently from the Horizon 2020 programme
\cite{h2020} had as their main goal to develop software solutions that provide a seamless and
transparent access by researchers to distributed computing and data resources.

@@

The EC projects considered in this paper are listed chronologically in Table~\ref{tab:eup}.
Each project paved the way for the following ones, according to a logic evolution in which
the role of the quality and reliability of the software has
consistently increased. As shown in Table~\ref{tab:feat}, the most recent projects are more
committed to, and aware, of software reliability practices.
Table~\ref{tab:feat} also shows the topics addressed in the following sections, highlighting
the commonalities and specific features on software reliability.

@@

The CI pipeline is composed from a set of required checks, their exit status
is reported back to GitHub. The changes can be prevented if those tests are
not successfully executed.
As each change is validated, the chances
of early detection of defects increase. Within this scenario, the cost of defect solving
is dramatically reduced and the reliability of the software solutions improved, as any
bug or design issue is likely to be detected and subsequently corrected in this phase.

@@

\subsubsection{Continuous Delivery}
As DevOps suggests, frequent releases positively affect the reliability of the software as
they reduce the time in which a given defect is exposed or has an impact, allowing the 
development teams to act promptly based on the regular feedback \cite{chen2015}. The software updates of
{\sl INDIGO--DataCloud} products, that have been taking place since the second major release, are passing through
a Continuous Delivery (CD) pipeline that adds the packaging of the software right after the
successful execution of the CI pipeline previously described. Consequently, CI--validated
software components are automatically delivered in online repositories. Nonetheless,
expert supervision to validate the results is required.
